FinalManuscript_DISCS_UG

### 2019-01-18 pcsc-paper/parts/02_01_paillier.tex
Added experiments
We have a mini-thesis na
Paper almost finished

\subsubsection{Key Generation}
We first define a function $L(x)$ as the largest integer $v$ greater than zero such that $x-1 \geq vn$.
We choose two large primes $p$ and $q$, and set $n = pq, \lambda = \mathrm{lcm}(p-1,q-1)$.
Then we select an integer $g$, $0\leq g \leq n^2$ such that $\mathrm{gcd}(L(g^\lambda \bmod n^2), n) = 1$.
We denote the public key as $(g,n)$ and the private key $(p,q)$.
\subsubsection{Encryption and Decryption}
The encryption function to encrypt a plaintext $m \in \mathbb{Z}_n$ given a public key $(g,n)$ is defined as
\begin{align*}
	E(m) = g^m \cdot r^n \mod{n^2},
\end{align*}
where $r$ is a random non-negative integer less than $n^2$.
The decryption function to decrypt a ciphertext $c \in \mathbb{Z}^*_{n^2}$ given a private key $(p,q)$ is defined as:
\begin{align*}
	D(c) = L(c^\lambda \bmod n^2) \times (L(g^\lambda \bmod n^2))^{-1} \mod n.
\end{align*}
\subsubsection{Homomorphic Properties of the Paillier Cryptosystem}
The Paillier cryptosystem supports additive homomorphism as well as the multiplication of a plaintext scalar to an encrypted message. These operations are defined as follows. These properties are used
For all $m_1,m_2 \in \mathbb{Z}_n$ and $k\in \mathbb{N}$, the following homomorphic properties hold.
\begin{align*}
	D(E(m_1)g^k\bmod n^2)=(m_1+k)\bmod n\\
	D(E(m_1)E(m_2)\bmod n^2)=(m_1+m_2)\bmod n\\
	D(E(m_1)^k\bmod n^2)= km_1\bmod n
\end{align*}

\subsection{Floating-Point Arithmetic}
The Paillier cryptosystem encrypts integers and allows operations on these integers. We now briefly go over several extensions to the Paillier cryptosystem discussed in the literature, which allow us to perform privacy-preserving floating-point arithmetic.

>>>>

\subsubsection{Key Generation}
We first define a function $L(x)$ as the largest integer $v$ greater than zero such that $x-1 \geq vn$.
We choose two large primes $p$ and $q$, and set $n = pq, \lambda = \mathrm{lcm}(p-1,q-1)$.
Then we select an integer $g$, $0\leq g \leq n^2$ such that $\mathrm{gcd}(L(g^\lambda \bmod n^2), n) = 1$.
We denote the public key as $(g,n)$ and the private key $(p,q)$.
\subsubsection{Encryption and Decryption}
The encryption function to encrypt a plaintext $m \in \mathbb{Z}_n$ given a public key $(g,n)$ is defined as
\begin{align*}
	E(m) = g^m \cdot r^n \mod{n^2},
\end{align*}
where $r$ is a random non-negative integer less than $n^2$.
The decryption function to decrypt a ciphertext $c \in \mathbb{Z}^*_{n^2}$ given a private key $(p,q)$ is defined as:
\begin{align*}
	D(c) = L\left(c^\lambda \bmod n^2\right) \cdot \left(L\left(g^\lambda \bmod n^2\right)\right)^{-1} \mod n.
\end{align*}
\subsubsection{Homomorphic Properties of the Paillier Cryptosystem}
The Paillier cryptosystem supports additive homomorphism as well as the multiplication of a plaintext scalar to an encrypted message. age. These homomorphic properties are used to add a plaintex scalar to a ciphertext, add two ciphertexts, and multiply a ciphertext by a plaintext scalar, respectively. They are defined as follows.

For all $m_1,m_2 \in \mathbb{Z}_n$ and $k\in \mathbb{N}$, the following homomorphic properties hold.
\begin{align*}
	D(E(m_1)g^k\bmod n^2)=(m_1+k)\bmod n\\
	D(E(m_1)E(m_2)\bmod n^2)=(m_1+m_2)\bmod n\\
	D(E(m_1)^k\bmod n^2)= km_1\bmod n
\end{align*}

\subsection{Floating-Point Arithmetic}
\label{sec:fp_operations}
The Paillier cryptosystem encrypts integers and allows operations on these integers. We now briefly go over several extensions to the Paillier cryptosystem discussed in the literature, which allow us to perform privacy-preserving floating-point arithmetic.

### 2019-01-18 pcsc-paper/parts/02_02_approximation.tex
Paper almost finished

\subsection{Known Approximations for Transcendental Functions}
Many methods to compute transcendental functions implemented in computer hardware and software libraries rely on direct knowledge of the function arguments. Several such approaches include using CORDIC (coordinate rotation digital computer), pseudo-division and pseudo-multiplication algorithms~\cite{walther_cordic_2000}, as well as using minimax polynomials, which are determined numerically using the Remez algorithm~\cite{harrison_computation_1999}. These algorithms, while efficient, rely on table lookups, which cannot be done in a privacy-preserving system where the function arguments cannot be directly determined.

We must therefore rely on other methods, such as using the Taylor series expansions of these functions. While also in common use, not all Taylor series expansions have a convenient radius of convergence. To circumvent this, we can rely on an approach presented by Khattri in~\cite{khattri_new_2009} to obtain closed form approximations to functions by approximating related integrals.

Khattri presents the following method to approximate the function $f(x)=\log(1+x)$.
Let $x = 1/n$ and consider the integral
\begin{align}
	\label{eq:log_integral}
	\int_{n}^{n+1}{\frac{1}{t}\diff t}=\log{\left(1+\frac{1}{n}\right)}.
\end{align}

This integral can be approximated using the five-point Gauss--Legendre quadrature rule~\cite{kythe_quadrature_2002}. We first convert the integral to an integral over the interval $[-1,1]$ using the following transformation:
\begin{align*}
	\int_a^b{f(x)\diff x}
	&= \frac{b-a}{2}\int_{-1}^{1}{f\left(\frac{b-a}{2}x+\frac{a+b}{2}\right)\diff x}.
\end{align*}
Then, we approximate the integral using the following summation:
\begin{align*}
	\int_{-1}^{1}{f(x)\diff x} &= \sum_{i=1}^{5}{w_if(x_i)},
\end{align*}
where
\begin{multicols}{2}
	\noindent
	\begin{align*}
		w_1 &= 0\\
		w_2 &= \frac{1}{21}\sqrt{245-14\sqrt{70}}\\
		w_3 &= -\frac{1}{21}\sqrt{245-14\sqrt{70}}\\
		w_4 &= \frac{1}{21}\sqrt{245+14\sqrt{70}}\\
		w_5 &= -\frac{1}{21}\sqrt{245+14\sqrt{70}}
	\end{align*}
	\columnbreak
	\begin{align*}
		x_1 &= \frac{128}{225}\\
		x_2 &= \frac{1}{900}\left( 322 + 13\sqrt{70}\right)\\
		x_3 &= \frac{1}{900}\left( 322 + 13\sqrt{70}\right)\\
		x_4 &= \frac{1}{900}\left( 322 - 13\sqrt{70}\right)\\
		x_5 &= \frac{1}{900}\left( 322 - 13\sqrt{70}\right)
	\end{align*}
\end{multicols}
Applying this procedure to the integral in Equation~\ref{eq:log_integral} yields the following approximation:
\begin{equation}\label{eq:standard_logarithm_quadrature}
	\log(1+x) =
	\frac{137x^5 + 2310x^4 + 9870x^3 + 15120x^2 + 7560x}
	{30x^5 + 900x^4 + 6300x^3 + 16800x^2 + 18900x + 7560}.
\end{equation}

While the closed form approximation presented by Khattri (Equation~\ref{eq:standard_logarithm_quadrature}) is accurate for values of $x$ near zero, it diverges from $\log{(1+x)}$ significantly for large values of $x$, as shown in Figure~\ref{fig:standard_logarithm_quadrature}.
\begin{figure}[!ht]
	\centering
	\begin{tikzpicture}
		\begin{axis}[
			% height=6cm, width=9cm,
			xmin=0, xmax=300,
			ymin=0, ymax=5.5, ytick={0,...,6},
			xlabel={$x$}, ylabel={$y$},
			grid=major,
			legend pos=south east,
			legend cell align=left,
		]
		\addplot[very thick, domain=0:300, samples=100]{1/30*(137*x^5 + 2310*x^4 + 9870*x^3 + 15120*x^2 + 7560*x)/(x^5 + 30*x^4 + 210*x^3 + 560*x^2 + 630*x + 252)};
		\addplot[densely dashed, domain=0:300, samples=100]{ln(x)};
		\legend{5-point quadrature approximation, Actual value of $\log(1+x)$}
		\end{axis}
	\end{tikzpicture}%
	\caption{Graph of $\log{(1+x)}$ and the approximation in Equation~\ref{eq:standard_logarithm_quadrature}}
	\label{fig:standard_logarithm_quadrature}
\end{figure}

However, the quadrature approach presented by Khattri allows us to approximate the logarithm in a larger range than is possible with the Taylor series, as the Taylor series of $\log(1+x)$ has an interval of convergence of $[-1,1]$.

>>>>

\subsection{Known Approximations for Transcendental Functions}
Many methods to compute transcendental functions implemented in computer hardware and software libraries rely on direct knowledge of the function arguments. Several such approaches include CORDIC (coordinate rotation digital computer), pseudo-division and pseudo-multiplication algorithms~\cite{walther_cordic_2000}, as well as minimax polynomials, which are determined numerically using the Remez algorithm~\cite{harrison_computation_1999}. These algorithms, while efficient, rely on table lookups, which cannot be done in a privacy-preserving system where the function arguments cannot be directly determined.

We must therefore rely on other methods, such as using the Taylor series expansions of these functions. While also in common use, not all Taylor series expansions have a convenient radius of convergence. To circumvent this, we can rely on an approach presented by Khattri in~\cite{khattri_new_2009} to obtain closed-form approximations to functions by approximating related integrals.

Khattri presents the following method to approximate the function $f(x)=\log\left(1+x\right)$.
Let $x = 1/n$ and consider the integral
\begin{align}
	\label{eq:log_integral}
	\int_{n}^{n+1}{\frac{1}{t}\diff t}=\log{\left(1+\frac{1}{n}\right)}.
\end{align}

This integral can be approximated using the five-point Gauss--Legendre quadrature rule~\cite{kythe_quadrature_2002}. We first convert the integral to an integral over the interval $[-1,1]$ using the following transformation:
\begin{align*}
	\int_a^b{f(x)\diff x}
	&= \frac{b-a}{2}\int_{-1}^{1}{f\left(\frac{b-a}{2}x+\frac{a+b}{2}\right)\diff x}.
\end{align*}
Then, we approximate the integral using the following summation:
\begin{align*}
	\int_{-1}^{1}{f\left(x\right)\diff x} &= \sum_{i=1}^{5}{w_i f\left(x_i\right)},
\end{align*}
where
\begin{multicols}{2}
	\noindent
	\begin{align*}
		w_1 &= 0\\
		w_2 &= \frac{1}{21}\sqrt{245-14\sqrt{70}}\\
		w_3 &= -\frac{1}{21}\sqrt{245-14\sqrt{70}}\\
		w_4 &= \frac{1}{21}\sqrt{245+14\sqrt{70}}\\
		w_5 &= -\frac{1}{21}\sqrt{245+14\sqrt{70}}
	\end{align*}
	\columnbreak
	\begin{align*}
		x_1 &= \frac{128}{225}\\
		x_2 &= \frac{1}{900}\left( 322 + 13\sqrt{70}\right)\\
		x_3 &= \frac{1}{900}\left( 322 + 13\sqrt{70}\right)\\
		x_4 &= \frac{1}{900}\left( 322 - 13\sqrt{70}\right)\\
		x_5 &= \frac{1}{900}\left( 322 - 13\sqrt{70}\right)
	\end{align*}
\end{multicols}
Applying this procedure to the integral in Equation~\ref{eq:log_integral} yields the following approximation:
\begin{equation}\label{eq:standard_logarithm_quadrature}
	\log\left(1+x\right) \approx
	\frac{137x^5 + 2310x^4 + 9870x^3 + 15120x^2 + 7560x}
	{30x^5 + 900x^4 + 6300x^3 + 16800x^2 + 18900x + 7560}.
\end{equation}

While the closed-form approximation presented by Khattri (Equation~\ref{eq:standard_logarithm_quadrature}) is accurate for values of $x$ near zero, it diverges from $\log{\left(1+x\right)}$ significantly for large values of $x$, as shown in Figure~\ref{fig:standard_logarithm_quadrature}.
\begin{figure}[!ht]
	\centering
	\begin{tikzpicture}
		\begin{axis}[
			% height=6cm, width=9cm,
			xmin=0, xmax=300,
			ymin=0, ymax=5.5, ytick={0,...,6},
			xlabel={$x$}, ylabel={$y$},
			grid=major,
			legend pos=south east,
			legend cell align=left,
		]
		\addplot[very thick, ACMDarkBlue, domain=0:300, samples=100]{1/30*(137*x^5 + 2310*x^4 + 9870*x^3 + 15120*x^2 + 7560*x)/(x^5 + 30*x^4 + 210*x^3 + 560*x^2 + 630*x + 252)};
		\addplot[densely dashed, domain=0:300, samples=100]{ln(x)};
		\legend{5-point quadrature approximation, Actual value of $\log(1+x)$}
		\end{axis}
	\end{tikzpicture}%
	\caption{Graph of $\log{\left(1+x\right)}$ and the approximation in Equation~\ref{eq:standard_logarithm_quadrature}}
	\label{fig:standard_logarithm_quadrature}
\end{figure}

However, the quadrature approach presented by Khattri allows us to approximate the logarithm in a larger range than is possible with the Taylor series, as the Taylor series of $\log\left(1+x\right)$ has an interval of convergence of $[-1,1]$.

### 2019-01-18 pcsc-paper/parts/03_exponentialandlogarithmic.tex
Paper almost finished

Applying the five-point Gauss--Legendre quadrature rule to approximate $\int_{n}^{\alpha n + \alpha}{\frac{1}{x}\diff x}$ with $\alpha = 1/16$ using SageMath 8.3, we arrive at the approximation:
\begin{align}\label{eq:optimal_log_approximation}
	\begin{split}
		&\log(1+x) \\
		&=\frac{137x^5 + 26685x^4 + 617370x^3 - 6498630x^2 - 121239315x - 257804775}
		{30(x^5 + 405x^4 + 27210x^3 + 488810x^2 + 2536005x + 3122577)}\\
		&+ \log{16}
	\end{split}
\end{align}

@@

\begin{figure}[ht]
	\centering
    \begin{tikzpicture}
		\begin{semilogyaxis}[
			% height=6cm, width=9cm,
			xmin=0, xmax=300,
			ymin=1e-10, ymax=1.4,
			xlabel={$x$}, ylabel={Absolute error},
			grid=major,
			legend pos=south east,
			legend cell align=left,
			tick scale binop=\times,
		]
		% \addplot[very thick, domain=0:300, samples=500]{abs(-1/30*(x^5*(30*ln(1/16) - 137) + 45*x^4*(270*ln(1/16) - 593) + 30*x^3*(27210*ln(1/16) - 20579) + 30*x^2*(488810*ln(1/16) + 216621) + 45*x*(1690670*ln(1/16) + 2694207) + 93677310*ln(1/16) + 257804775)/(x^5 + 405*x^4 + 27210*x^3 + 488810*x^2 + 2536005*x + 3122577) - ln(x + 1))};
		% \addplot[very thick, domain=0:300, samples=300]{abs(1/30*(137*x^5 + 26685*x^4 + 617370*x^3 - 6498630*x^2 - 121239315*x - 257804775)/(x^5 + 405*x^4 + 27210*x^3 + 488810*x^2 + 2536005*x + 3122577) - ln(1/16) - ln(x + 1))};
		\addplot[very thick] table[x=x, y=error] {plots/abs_err_scaledlog.txt};
		\addplot[densely dashed, domain=0:300, samples=100]{abs(1/30*(137*x^5 + 2310*x^4 + 9870*x^3 + 15120*x^2 + 7560*x)/(x^5 + 30*x^4 + 210*x^3 + 560*x^2 + 630*x + 252) - ln(x + 1))};
		\legend{scaled ($\alpha = 1/16$), original ($\alpha = 1$)}
	\end{semilogyaxis}
	\end{tikzpicture}%
	\caption{Comparison of the absolute error between the original approximation in Equation~\ref{eq:standard_logarithm_quadrature} and the scaled approximation in Equation~\ref{eq:optimal_log_approximation}}
	\label{fig:log_error_comparison}
\end{figure}

@@

It can be determined numerically using SageMath 8.3 that a local minimum of $\max(A(0,\alpha),A(255,\alpha))$ on the interval $[0,1]$ is at $\alpha = 1/16$, which is found at the intersection of the two graphs.
\begin{figure}[!ht]
	\centering
    \begin{tikzpicture}
		\begin{axis}[
			% height=6cm, width=9cm,
			xmin=0, xmax=1,
			ymin=0, ymax=1.3,
			xlabel={$\alpha$}, ylabel={Absolute error},
			grid=major,
			legend pos=south east,
			legend cell align=left,
			tick scale binop=\times,
		]
		\addplot[ACMDarkBlue, domain=0:1, samples=200]{1/30*abs((137*x^5 + 1625*x^4 + 2000*x^3 - 2000*x^2 - 30*(x^5 + 25*x^4 + 100*x^3 + 100*x^2 + 25*x + 1)*ln(x) - 1625*x - 137)/(x^5 + 25*x^4 + 100*x^3 + 100*x^2 + 25*x + 1))};
		\addplot[ACMOrange, domain=0:1, samples=200]{abs(-1/30*(32346117281250*x^5*ln(x) - 150633093005312*x^5 - 6979321856000*x^4 - 33554432000*x^3 + 131072000*x^2 + 30*(x^5 + 25*x^4 + 100*x^3 + 100*x^2 + 25*x + 1)*ln(x) + 38250*(x^5 + 20*x^4 + 60*x^3 + 40*x^2 + 5*x)*ln(x) + 19507500*(x^5 + 15*x^4 + 30*x^3 + 10*x^2)*ln(x) + 4974412500*(x^5 + 10*x^4 + 10*x^3)*ln(x) + 634237593750*(x^5 + 5*x^4)*ln(x) + 416000*x + 137)/(1099511627776*x^5 + 107374182400*x^4 + 1677721600*x^3 + 6553600*x^2 + 6400*x + 1) - 8*ln(2))};
		\legend{$x=0$, $x=255$}
	\end{axis}
	\end{tikzpicture}%
	\caption{Graphs of $A(x,\alpha)$ at $x=0$ and $x=255$}
	\label{fig:endpoint_plot}
\end{figure}
We then consider the function $A(x,1/16)$, shown in Figure~\ref{fig:single_alpha_plot}.
It was determined numerically using SageMath 8.3 that the local maxima of $A(x,1/16)$ on the interval occur at $x=0$ and $x=255$. Since a scaling factor other than $\alpha=1/16$ yields a higher absolute error at either $x=0$ or $x=255$, as seen in Figure~\ref{fig:endpoint_plot}, it follows that a scaling factor of $\alpha = 1/16$ minimizes the maximum absolute error of the scaled approximation for $\log{(1+x)}$, within the range $x \in [0, 255]$.
\begin{figure}[ht]
	\centering
	\begin{tikzpicture}
		\begin{axis}[
			xmin=0, xmax=300,
			ymin=0, ymax=0.035,
			xlabel={$x$}, ylabel={Absolute error},
			grid=major,
			legend pos=south east,
			legend cell align=left,
			tick scale binop=\times,
		]
		% \addplot[very thick, domain=0:300, samples=300]{abs(1/30*(137*x^5 + 26685*x^4 + 617370*x^3 - 6498630*x^2 - 121239315*x - 257804775)/(x^5 + 405*x^4 + 27210*x^3 + 488810*x^2 + 2536005*x + 3122577) - ln(1/16) - ln(x + 1))};
		\addplot[very thick] table[x=x, y=error] {plots/abs_err_scaledlog.txt};
	\end{axis}
	\end{tikzpicture}%
	\caption{Graph of $A(x,1/16)$, the absolute error of the approximation in Equation~\ref{eq:optimal_log_approximation} given a scaling factor of $\alpha=1/16$}
	\label{fig:single_alpha_plot}
\end{figure}

>>>>

Applying the five-point Gauss--Legendre quadrature rule to approximate $\int_{n}^{\alpha n + \alpha}{\frac{1}{x}\diff x}$ with $\alpha = 1/16$ using SageMath 8.3, we arrive at the approximation:
\begin{align}\label{eq:optimal_log_approximation}
	\begin{split}
		&\log(1+x) \\
		&\approx \frac{137x^5 + 26685x^4 + 617370x^3 - 6498630x^2 - 121239315x - 257804775}
		{30(x^5 + 405x^4 + 27210x^3 + 488810x^2 + 2536005x + 3122577)}\\
		&+ \log{16}
	\end{split}
\end{align}

@@

\begin{figure}[ht]
	\centering
    \begin{tikzpicture}
		\begin{semilogyaxis}[
			% height=6cm, width=9cm,
			xmin=0, xmax=300,
			ymin=1e-10, ymax=1.4,
			xlabel={$x$}, ylabel={Absolute error},
			grid=major,
			legend pos=south east,
			legend cell align=left,
			tick scale binop=\times,
		]
		% \addplot[very thick, domain=0:300, samples=500]{abs(-1/30*(x^5*(30*ln(1/16) - 137) + 45*x^4*(270*ln(1/16) - 593) + 30*x^3*(27210*ln(1/16) - 20579) + 30*x^2*(488810*ln(1/16) + 216621) + 45*x*(1690670*ln(1/16) + 2694207) + 93677310*ln(1/16) + 257804775)/(x^5 + 405*x^4 + 27210*x^3 + 488810*x^2 + 2536005*x + 3122577) - ln(x + 1))};
		% \addplot[very thick, domain=0:300, samples=300]{abs(1/30*(137*x^5 + 26685*x^4 + 617370*x^3 - 6498630*x^2 - 121239315*x - 257804775)/(x^5 + 405*x^4 + 27210*x^3 + 488810*x^2 + 2536005*x + 3122577) - ln(1/16) - ln(x + 1))};
		\addplot[very thick, ACMDarkBlue] table[x=x, y=error] {plots/abs_err_scaledlog.txt};
		\addplot[densely dashed, domain=0:300, samples=100]{abs(1/30*(137*x^5 + 2310*x^4 + 9870*x^3 + 15120*x^2 + 7560*x)/(x^5 + 30*x^4 + 210*x^3 + 560*x^2 + 630*x + 252) - ln(x + 1))};
		\legend{Scaled quadrature ($\alpha = 1/16$), Original quadrature ($\alpha = 1$)}
	\end{semilogyaxis}
	\end{tikzpicture}%
	\caption{Comparison of the absolute error between the original approximation in Equation~\ref{eq:standard_logarithm_quadrature} and the scaled approximation in Equation~\ref{eq:optimal_log_approximation}}
	\label{fig:log_error_comparison}
\end{figure}

@@

It can be determined numerically using SageMath 8.3 that a local minimum of $\max(A(0,\alpha),A(255,\alpha))$ on the interval $[0,1]$ is at $\alpha = 1/16$, which is found at the intersection of the two graphs.
\begin{figure}[!ht]
	\centering
    \begin{tikzpicture}
		\begin{semilogyaxis}[
			% height=6cm, width=9cm,
			xmin=0, xmax=1,
			ymin=5e-7, ymax=1.3,
			xlabel={$\alpha$}, ylabel={Absolute error},
			grid=major,
			extra x ticks={1/16},
			extra x tick labels={$\frac{1}{16}$},
			legend pos=south east,
			legend cell align=left,
			tick scale binop=\times,
		]
		% \addplot[thick, ACMDarkBlue, domain=0:1, samples=200]{1/30*abs((137*x^5 + 1625*x^4 + 2000*x^3 - 2000*x^2 - 30*(x^5 + 25*x^4 + 100*x^3 + 100*x^2 + 25*x + 1)*ln(x) - 1625*x - 137)/(x^5 + 25*x^4 + 100*x^3 + 100*x^2 + 25*x + 1))};
		% \addplot[thick, ACMOrange, domain=0:1, samples=200]{abs(-1/30*(32346117281250*x^5*ln(x) - 150633093005312*x^5 - 6979321856000*x^4 - 33554432000*x^3 + 131072000*x^2 + 30*(x^5 + 25*x^4 + 100*x^3 + 100*x^2 + 25*x + 1)*ln(x) + 38250*(x^5 + 20*x^4 + 60*x^3 + 40*x^2 + 5*x)*ln(x) + 19507500*(x^5 + 15*x^4 + 30*x^3 + 10*x^2)*ln(x) + 4974412500*(x^5 + 10*x^4 + 10*x^3)*ln(x) + 634237593750*(x^5 + 5*x^4)*ln(x) + 416000*x + 137)/(1099511627776*x^5 + 107374182400*x^4 + 1677721600*x^3 + 6553600*x^2 + 6400*x + 1) - 8*ln(2))};
		\addplot[thick, ACMDarkBlue, domain=0:1] table[x=alpha, y=x0] {plots/abs_err_sl_0_255.txt};
		\addplot[thick, ACMOrange, domain=0:1] table[x=alpha, y=x255] {plots/abs_err_sl_0_255.txt};
		\legend{$x=0$, $x=255$}
	\end{semilogyaxis}
	\end{tikzpicture}%
	\caption{Graphs of $A(x,\alpha)$ at $x=0$ and $x=255$}
	\label{fig:endpoint_plot}
\end{figure}
We then consider the function $A(x,1/16)$, shown in Figure~\ref{fig:single_alpha_plot}.
It was determined numerically that the local maxima of $A(x,1/16)$ on the interval occur at $x=0$ and $x=255$. Since a scaling factor other than $\alpha=1/16$ yields a higher absolute error at either $x=0$ or $x=255$, as seen in Figure~\ref{fig:endpoint_plot}, it follows that a scaling factor of $\alpha = 1/16$ minimizes the maximum absolute error of the scaled approximation for $\log{(1+x)}$, within the range $x \in [0, 255]$.
\begin{figure}[ht]
	\centering
	\begin{tikzpicture}
		\begin{semilogyaxis}[
			xmin=0, xmax=300,
			ymin=1e-15, ymax=0.035,
			xlabel={$x$}, ylabel={Absolute error},
			grid=major,
			legend pos=south east,
			legend cell align=left,
			tick scale binop=\times,
		]
		% \addplot[very thick, domain=0:300, samples=300]{abs(1/30*(137*x^5 + 26685*x^4 + 617370*x^3 - 6498630*x^2 - 121239315*x - 257804775)/(x^5 + 405*x^4 + 27210*x^3 + 488810*x^2 + 2536005*x + 3122577) - ln(1/16) - ln(x + 1))};
		\addplot[thick, ACMDarkBlue] table[x=x, y=error] {plots/abs_err_scaledlog.txt};
	\end{semilogyaxis}
	\end{tikzpicture}%
	\caption{Graph of $A(x,1/16)$, the absolute error of the approximation in Equation~\ref{eq:optimal_log_approximation} given a scaling factor of $\alpha=1/16$}
	\label{fig:single_alpha_plot}
\end{figure}

### 2019-01-18 pcsc-paper/parts/04_inversetrig.tex
Paper almost finished

To obtain a Gauss--Legendre quadrature-based approximation, we take the following integral,
\begin{align*}
	\arctan x = \int_0^x{\frac{1}{1+x^2}\diff x},
\end{align*}
and perform five-point Gauss--Legendre quadrature to obtain the following closed-form approximation:
\begin{align} \label{eq:arctan_quadrature}
	\begin{split}
		&\arctan x \\
		&= \frac
		{4\left(225x^9 + 15925x^7 + 144753x^5 + 350595x^3 + 238140x\right)}
		{15\left(x^{10} + 480x^8 + 11760x^6 + 64120x^4 + 114660x^2 + 63504\right)}
	\end{split}
\end{align}

We note that in order to evaluate the quadrature approximation in Equation~\ref{eq:arctan_quadrature}, we must compute the value of $x^{10}$, which requires several rounds of the squaring and multiplication protocols to compute. We now compare the quadrature approximation to the partial sum composed of the first 4 terms of Euler's series expansion in Equation~\ref{eq:arctan_euler}, which requires a similar number of floating-point operations to compute:
\begin{align} \label{eq:arctan_euler_partial}
	\begin{split}
		&\sum_{n=0}^{3}
		{
		\frac
			{2^{2n}(n!)^2}
			{(2n+1)!}
		\frac
			{x^{2n+1}}
			{(1+x^2)^{n+1}}
		}\\
		&= \frac
		{965x^9 + 2370x^7 + 2688x^5 + 1470x^3 + 315x}
		{315\left(x^{10} + 5x^8 + 10x^6 + 10x^4 + 5x^2 + 1\right)}
	\end{split}
\end{align}

Figure~\ref{fig:arctan_error} shows that despite requiring a similar number of floating-point operations, the quadrature based approximation in Equation~\ref{eq:arctan_quadrature} gives accurate approximations over a larger interval, making it a more viable candidate for efficient computation.
\begin{figure}[!ht]
		\centering
		\includegraphics[width=.9\linewidth]{figures/single_alpha_plot.png}
		\caption{Comparison of the absolute error of the approximations for $\arctan x$ using quadrature approximation (Equation~\ref{eq:arctan_quadrature}) and the first 4 terms of Euler's series (Equation~\ref{eq:arctan_euler_partial})}
		\label{fig:arctan_error}
\end{figure}

@@

To carry out the conditional check required to use these identities, a privacy-preserving comparison protocol may be used, such as in~\cite{veugen_improving_2012}.
\begin{figure}[!ht]
		\centering
		\includegraphics[width=.9\linewidth]{figures/arctan_error_2.png}
		\caption{Absolute error of the approximation for $\arctan x$ using quadrature approximation (Equation~\ref{eq:arctan_quadrature}) over the interval $[-1,1]$}
		\label{fig:arctan_error_2}
\end{figure}

>>>>

To obtain a Gauss--Legendre quadrature-based approximation, we take the following integral,
\begin{align*}
	\arctan x = \int_0^x{\frac{1}{1+t^2}\diff t},
\end{align*}
and perform five-point Gauss--Legendre quadrature to obtain the following closed-form approximation:
\begin{align} \label{eq:arctan_quadrature}
	\begin{split}
		&\arctan x \\
		&\approx \frac
		{4\left(225x^9 + 15925x^7 + 144753x^5 + 350595x^3 + 238140x\right)}
		{15\left(x^{10} + 480x^8 + 11760x^6 + 64120x^4 + 114660x^2 + 63504\right)}
	\end{split}
\end{align}

We note that in order to evaluate the quadrature approximation in Equation~\ref{eq:arctan_quadrature}, we must compute the value of $x^{10}$, which requires several rounds of the squaring and multiplication protocols to compute. We now compare the quadrature approximation to the partial sum composed of the first five terms of Euler's series expansion in Equation~\ref{eq:arctan_euler}, which requires a similar number of floating-point operations to compute:
\begin{align} \label{eq:arctan_euler_partial}
	\begin{split}
		&\sum_{n=0}^{4}
		{
		\frac
			{2^{2n}(n!)^2}
			{(2n+1)!}
		\frac
			{x^{2n+1}}
			{(1+x^2)^{n+1}}
		}\\
		&= \frac
		{965x^9 + 2370x^7 + 2688x^5 + 1470x^3 + 315x}
		{315\left(x^{10} + 5x^8 + 10x^6 + 10x^4 + 5x^2 + 1\right)}
	\end{split}
\end{align}

Figure~\ref{fig:arctan_error} shows that despite requiring a similar number of floating-point operations, the quadrature-based approximation in Equation~\ref{eq:arctan_quadrature} gives accurate approximations over a larger interval, making it a more viable candidate for efficient computation as seen in Figure~\ref{fig:arctan_comparison}.
\begin{figure}[!ht]
	\centering
	\begin{tikzpicture}
		\begin{semilogyaxis}[
			xmin=-40, xmax=40,
			ymin=1e-11, ymax=2,
			xlabel={$x$}, ylabel={Absolute error},
			grid=major,
			legend pos=south east,
			legend cell align=left,
			tick scale binop=\times,
		]
		\addplot[thick, ACMDarkBlue] table[x=x, y=gl] {plots/abs_err_arctan.txt};
		\addplot[thick, ACMOrange] table[x=x, y=euler] {plots/abs_err_arctan.txt};
		\legend{Gauss--Legendre quadrature, Euler's series}
		\end{semilogyaxis}
	\end{tikzpicture}%
	\caption{Comparison of the absolute error of the approximations for $\arctan x$ using quadrature approximation (Equation~\ref{eq:arctan_quadrature}) and the first five terms of Euler's series (Equation~\ref{eq:arctan_euler_partial})}
	\label{fig:arctan_error}
\end{figure}

\begin{figure}[ht]
	\centering
    \begin{tikzpicture}
		\begin{axis}[
			xmin=-12, xmax=12,
			ymin=-2.6, ymax=2.6,
			xlabel={$x$}, ylabel={$y$},
			grid=major,
			legend pos=south east,
			legend cell align=left,
			tick scale binop=\times,
		]
		\addplot[very thick, ACMDarkBlue] table[x=x, y=yhat] {plots/plot_arctan.txt};
		\addplot[densely dashed] table[x=x, y=y] {plots/plot_arctan.txt};
		\legend{5-point quadrature approximation, Actual value of $\arctan x$}
	\end{axis}
	\end{tikzpicture}%
	\caption{Graph of $\arctan x$ and the approximation in Equation~\ref{eq:arctan_quadrature}}
	\label{fig:arctan_comparison}
\end{figure}

@@

To carry out the conditional check required to use these identities, a privacy-preserving comparison protocol may be used, such as in~\cite{veugen_improving_2012}.
\begin{figure}[!ht]
	\centering
	\begin{tikzpicture}
		\begin{semilogyaxis}[
			xmin=-1.25, xmax=1.25,
			ymin=1e-18, ymax=5e-5, ytickten={-18,-16,...,0},
			xlabel={$x$}, ylabel={Absolute error},
			grid=major,
			legend pos=south east,
			legend cell align=left,
			tick scale binop=\times,
		]
		\addplot[thick, ACMDarkBlue] table[x=x, y=gl] {plots/abs_err_arctan.txt};
		\end{semilogyaxis}
	\end{tikzpicture}%
	\caption{Absolute error of the approximation for $\arctan x$ using quadrature approximation (Equation~\ref{eq:arctan_quadrature}) over the interval $[-1,1]$}
	\label{fig:arctan_error_2}
\end{figure}
